---
title: "Homework #1"
author: "Taeyoung Heo(UMID # 2914 5801)"
format: 
 html:
  toc: false
---

## Problem1 - Wine data

<b>P1-a.</b> import "wine.data" into a data.frame in R

```{r}
winedata <- read.table("wine.data", sep = ",")
colnames(winedata) <- c("class", "alcohol", "acid", "ash", "alcalinity", "magnesium", "phenols", "flavanoids", "nonflavanoid_phenols", "proanthocyanins", "color_intensity", "hue", "protein", "proline")        
```
<br>
<b>P1-b.</b> check the number of wines within each class <br> [class1 = 59, class2 = 71, class3 = 48]

```{r}
wineclass <- winedata$class
count <- c(length(which(wineclass==1)), length(which(wineclass == 2)), length(which(wineclass == 3)))
report <- c(59, 71, 48)

if(count[1] == report[1] & count[2] == report[2] & count[3] == report[3]) { 
  print("correct as reported in wine.names")
  } else { 
  print("wrong as reported in wine.names")
  }
```
<br>
<b>P1-c-1.</b> the correlation between alcohol content and color intensity <br> the correlation : <b>0.5463642</b>

```{r}
alcohol <- winedata$alcohol
color <- winedata$color_intensity
cor(alcohol, color)
```
<br>
<b>P1-c-2.</b> the highest and the lowest correlation <br>
the highest : <b>class1(0.4082913)</b> <br>
the lowest : <b>class2(0.2697891)</b> 

```{r}
class <- winedata$class

alcohol1 <- winedata$alcohol[winedata$class == 1]
alcohol2 <- winedata$alcohol[winedata$class == 2]
alcohol3 <- winedata$alcohol[winedata$class == 3]

color1 <- winedata$color[winedata$class == 1]
color2 <- winedata$color[winedata$class == 2]
color3 <- winedata$color[winedata$class == 3]

gcor <- c(cor(alcohol1, color1), cor(alcohol2, color2), cor(alcohol3, color3))
gcor 
```
<br>
<b>P1-c-3.</b> the alcohol content of the wine with the highest color intensity <br>
the alcohol content with the hightest color intensity : <b>14.34</b>

```{r}
length(which(color == max(color)))
alcohol_hcolor <- winedata$alcohol[winedata$color_intensity == max(color)]
alcohol_hcolor
```
<br>
<b>P1-c-4.</b> the percentage of the wines with a higher content of proanthocyanins compare to ash <br>
the percentage : <b>8.426966%</b>

```{r}
proanthocyanins <- winedata$proanthocyanins
proan_higher <- winedata$proanthocyanins[winedata$proanthocyanins > winedata$ash]
length(proan_higher) / length(proanthocyanins) * 100
```
<br>
<b>P1-d.</b> table for the average value of each variable <br>
the table : <b>average_table</b>

```{r}
average_table <- matrix(0, nrow = 4, ncol = length(colnames(winedata))-1)
colnames(average_table) <- colnames(winedata[2:14])
rownames(average_table) <- c("overall", "class1", "class2", "class3")

cal_mat <- as.matrix(winedata)
n <- ncol(cal_mat)-1

ave_overall <- numeric(n)
ave_class1 <- numeric(n)
ave_class2 <- numeric(n)
ave_class3 <- numeric(n)

for ( i in 1:n) {
  ave_overall[i] <- mean(cal_mat[,i+1])
  ave_class1[i] <- mean(cal_mat[cal_mat[,1] == 1, i+1])
  ave_class2[i] <- mean(cal_mat[cal_mat[,1] == 2, i+1])
  ave_class3[i] <- mean(cal_mat[cal_mat[,1] == 3, i+1])
}

average_table["overall",] <- ave_overall
average_table["class1",] <- ave_class1
average_table["class2",] <- ave_class2
average_table["class3",] <- ave_class3

average_table
```
<br>
<b>P1-e.</b> t-test to examine whether the level of phenols differs across the three classes <br>
<br>
We can reject the hypothesis that the level of phenols is the same across the three classes because all three p-values are less than 0.01. So we can conclude that the level of phenols differs across the three classes. <br>
<br>
t-test p-value between class1 and class2 : <b>1.889e-11</b> <br>
t-test p-value between class1 and class3 : <b>2.2e-16</b> <br>
t-test p-value between class2 and class3 : <b>1.622e-10</b>

```{r}
phenols1 <- winedata$phenols[winedata$class == 1]
phenols2 <- winedata$phenols[winedata$class == 2]
phenols3 <- winedata$phenols[winedata$class == 3]

t.test(phenols1, phenols2)
t.test(phenols1, phenols3)
t.test(phenols2, phenols3)
```

## Problem2 - AskAManager.org Data 
<b>P2-a.</b> import AskAManager.csv into a data.frame

```{r}
salary <- read.csv("AskAManager.csv")
```
<br>
<b>P2-b.</b> simplify the variable names
```{r}
colnames(salary) <- c("number", "time", "age", "industry", "title", "title_context", "salary", "bonus", "currency", "currency_other", "income_context", "country", "state", "city", "experience", "experience_field", "education", "gender", "race")
```
<br>
<b>P2-c.</b> restrict the data to those paid in USD <br>
the number of observations before : <b>28,062</b> <br>
the number of oberservations after the restriction : <b>23,374</b> 
```{r}
length(salary$currency)
length(salary$currency[salary$currency == "USD"])
salary_res <- salary[salary$currency == "USD", ]

nrow(salary_res) == length(salary$currency[salary$currency == "USD"])
```
<br>
<b>P2-d.</b> Eliminate any rows for which their age, years of experience in their field, and years of experience total are impossible <br>
the number of observation after the elimination : <b>23,116</b>
```{r}
age <- salary_res$age
field <- salary_res$experience_field
total <- salary_res$experience

age[!duplicated(age)]
field[!duplicated(field)]
total[!duplicated(total)]

age_kind <- c("25-34" = 16, "45-54" = 36, "35-44" = 26, "18-24" = 6, "65 or over" = 56, "55-64" = 46, "under 18" = 0)

experience_kind <- c("5-7 years" = 5, "2 - 4 years" = 2, "8 - 10 years" = 8, "21 - 30 years" = 21, "11 - 20 years" = 11, "41 years or more" = 41, "31 - 40 years" = 31, "1 year or less" = 1 )

age_trans <- age_kind[age]
field_trans <- experience_kind[field]
total_trans <- experience_kind[total]
n_usd <- nrow(salary_res)
memo <- logical(n_usd)

for(i in 1:n_usd) { 
  # if age < 18, eliminate the data
  # if age - 18 < field experience, eliminate the data
  # if age - 18 < total experience, eliminate the data
  # if total experience < field experience, eliminate the data
  
  if(age_trans[i] == 0) {
       memo[i] = FALSE
  }
  else if(field_trans[i]>age_trans[i]) {
       memo[i] = FALSE
  }
  else if(total_trans[i]>age_trans[i]) {
       memo[i] = FALSE
  } 
  else if(field_trans[i]>total_trans[i]) {
       memo[i] = FALSE
  }
  else {
       memo[i] = TRUE
  }
}

salary_year <- salary_res[memo, ]

salary_year$age[!duplicated(salary_year$age)] #test whether age is under 18 

#test example for whether their is reasonable with total experience 
experience_test <- salary_year$age[salary_year$experience == "41 years or more"]
experience_test[!duplicated(experience_test)] 

#test example for whether their is reasonable with field experience 
fieldexp_test <- salary_year$age[salary_year$experience_field == "41 years or more"]
fieldexp_test[!duplicated(fieldexp_test)] 

#test example for whether filed experience is less than total experience
exp_fieldexp_test <- salary_year$experience_field[salary_year$experience == "8 - 10 years"]
exp_fieldexp_test[!duplicated(exp_fieldexp_test)] 

nrow(salary_year)
```
<br>
<b>P2-e.</b> eliminate any rows with extremely low or high salaries. <br>
<br>
The other monetary compensation is added to salalry in order to eliminate false slalrries.<br>
<br>
The extreme high salary : <b>1459608 USD</b> (mean + 2*standard deviation) <br>
-> There are values on the right side of the distribution that differ from the mean by more than twice the standard deviation. <br>
<br>
The extreme low salary : <b>1,000 USD</b> <br>
-> There's a very heavy tail on the left side of the distribution.<br>
<br>
The sample size : <b>23044</b>

```{r}
income <- salary_year$salary
bonus <- salary_year$bonus
bonus[is.na(bonus)] <- 0
monetary_reward <- income + bonus
summary(monetary_reward)

n_year <- length(monetary_reward)

mean_monetary <- mean(monetary_reward)
sd_monetary <- sd(monetary_reward)
mean_monetary + 2*sd_monetary
mean_monetary - sd_monetary

hist(monetary_reward[monetary_reward <= 1459608 & monetary_reward >= 1000], breaks = 100, main = "Monetary reward between 1,000 and 1,459,608 USD", xlab = "monetary reward", ylab = "number of people") 
hist(monetary_reward[monetary_reward > 1459608], breaks = 100, main = "Monetary reward over 1,459,608 USD", xlab = "monetary reward", ylab = "number of people") 
hist(monetary_reward[monetary_reward < 1000 & monetary_reward > 0], breaks = 100, main = "Monetary reward less than 1,000 USD", xlab = "monetary reward", ylab = "number of people")

salary_year$monetary_reward <- monetary_reward
salary_monetary <- salary_year[monetary_reward <= 1459608 & monetary_reward >= 1000, ]
nrow(salary_monetary)
```

## Problem 3 - Palindromic Numbers

<b>P3-a.</b> function "isPalindromic" that check if a given positive integer is a palindromic 

```{r}
#' Palindromic of Integers 
#'
#' Determine whether a positive intger is palindromic or not 
#'
#' This function takes a positive integer x and returns a logical value whether the integer is a palindromic number or not and return reversed number to show that whether the reversed is same with the original number.
#'
#' @param x a positive integer 
#' @param na.rm logical, should NA's be removed? Default is FALSE
#'
#' @return list of a logical value and a numeric value
#' @export
#'
#' @examples
#' x <- 343
#' isPalindromic(x)
#' $isPalindromic
#' [1] TRUE
#' $reversed
#' [1] "343"
isPalindromic <- function(x, na.rm = FALSE) {
  if(!is.numeric(x)) { 
    # this code is from STATS 506 class note
    warning("x must be a positive integer, attemting to convert")
    suppressWarnings(x <- as.numeric(x))
    if(all(is.na(x))) {
      stop("x must be a positive integer or convertible to a positive integer")
    }
  }
  if(length(x) == 0) {
    # this code is from STATS 506 class note
    stop("x must have strictly positive length")
  }
  if(!is.logical(na.rm)) {
    # this code is from STATS 506 class note
    warning("na.rm must be logical")
  }
  if(x <= 0 ) {
    stop("x must be a positive integer or convertible to a positive integer")
  }
  if(x != round(x)) {
    stop("x must be a positive integer or convertible to a positive integer")
  }
  
  y <- as.character(x)
  z <- as.numeric(unlist(strsplit(y, "")))
  k <- length(z)
  check_palindromic <- logical(k)
  reverse <- numeric(k)
  for(i in 1:k) {
    if(z[i] == z[k+1-i]) {
      check_palindromic[i] <- TRUE
    } else { 
      check_palindromic[i] <- FALSE
    }
    reverse[i] <- z[k+1-i]
  }
  reverse_return <- as.numeric(paste(reverse, collapse = ""))
  if(all(check_palindromic == TRUE)) {
    return(list(isPalindromic = TRUE, reversed = reverse_return))
  }
  else { 
    return(list(isPalindromic = FALSE, reversed = reverse_return))
  }
}

isPalindromic(1100) # check the "isPalindromic()" function 
```
<br>
<b>P3-b.</b> create a nextPalindrome function that finds the next palindromic number strictly greater than the input

```{r}
#' Next Palindrome
#'
#' Find the next palindromic number strictly greater than the input
#' 
#' This function takes a positive integer x and check from x+1 to the next paliindromic number whether the integer is a palindromic number or not and return the next palindromic number. 
#'
#' @param x a positive integer
#' @param na.rm logical, should NA's be removed? Default is FALSE
#'
#' @return a numeric value
#' @export
#'
#' @examples
#' x <- 1523
#' nextPalindrome(x)
#' [1] "1551"
nextPalindrome <- function(x, na.rm = FALSE) {
  if(!is.numeric(x)) {
    # this code is from STATS 506 class note
    warning("x must be a positive integer, attemting to convert")
    suppressWarnings(x <- as.numeric(x))
    if(all(is.na(x))) {
      stop("x must be a positive integer or convertible to a positive integer")
    }
  }
  if(length(x) == 0) {
    # this code is from STATS 506 class note
    stop("x must have strictly positive length")
  }
  if(!is.logical(na.rm)) {
    # this code is from STATS 506 class note
    warning("na.rm must be logical")
  }
  if(x <= 0 ) {
    stop("x must be a positive integer or convertible to a positive integer")
  }
  if(x != round(x)) {
    stop("x must be a positive integer or convertible to a positive integer")
  }
  
  y <- as.character(x+1)
  z <- as.numeric(unlist(strsplit(y, "")))
  k <- length(z)
  check <- logical(k)
  next_palindrome <- numeric(k)
  while(!all(check == TRUE)) {
    for(i in 1:k) {
      if(z[i] == z[k+1-i]) {
        check[i] <- TRUE
      } else { 
        check[i] <- FALSE
      }
      next_palindrome[i] <- z[i]
    }
    y <- as.character(as.numeric(y)+1)
    z <- as.numeric(unlist(strsplit(y, "")))
    k <- length(z)
  }
  next_palindrome <- as.numeric(paste(next_palindrome, collapse = ""))
  return(nextPalindrome = next_palindrome)
  
}

nextPalindrome(1523) # check the "nextPalindrome()" function
```
<br>
<b>P3-c.</b> Use above functions to find the next palindrome for each of the following: (391, 9928, 19272719, 109, 2) <br>
<br>
i. 391 : <b>393</b> <br>
ii. 9928 : <b>9999</b> <br>
iii. 19272719 : <b>19277291</b> <br>
iv. 109 : <b>111</b> <br>
v. 2 : <b>3</b> 

```{r}
nextPalindrome(391)
nextPalindrome(9928)
nextPalindrome(19272719)
nextPalindrome(109)
nextPalindrome(2)
```
